Missing data: Current practice in football research and recommendations for reporting missingness

### 2020-09-03 paper/paper.Rmd

Missing data are values that should have been observed, but were not. Since the
values are unobserved, this can mean the population of interest might not be 
properly sampled, inducing bias into study results. Left undetected or poorly handled, missing data can undermine the validity and accuracy of research results
[@Sainani2015; @Nakagawa2008; @Sterne2009]. Data can go
missing many ways. For example, accidentally skipping a survey question,
equipment failure, or intentionally not recording values. Before
analysis, missing data must be handled, with values typically deleted or
imputed. There is no universal approach to handling missing data. 
Contextual factors--such as, the study design and objective, and
pattern of missingness--determine how missings should be handled, on a
case-by-case basis [@Sainani2015].
The effect of missing data on the validity and accuracy of a study's
results is largely determined by the reasons for missingness, and how
missing data are handled.

A missing value should have been observed or measured in the data, but was not. For example, a goal scored in football takes place, but the scorekeeper wasn't paying attention and did not record an event. This "goal" is now a missing value. Compare this to a game being rained out and no game taking place. There is no missing data because there is no data to be observed. A final example - a study screens out individuals who do not meet criteria for the study. These are not missing values, because there was no intent for these to be observed or measured.

>>>>

Missing data are values that should have been observed, but were not. Since the
values are unobserved, this can mean the population of interest might not be 
properly sampled, inducing bias into study results. Left undetected or poorly handled, missing data can undermine the validity and accuracy of research results
[@Sainani2015; @Nakagawa2008; @Sterne2009]. Data can go
missing many ways. For example, accidentally skipping a survey question,
equipment failure, or intentionally not recording values. Before
analysis, missing data must be handled, with values typically deleted or
imputed. There is no universal approach to handling missing data. 
Contextual factors--such as, the study design and objective, and
pattern of missingness--determine how missings should be handled, on a
case-by-case basis [@Sainani2015].

For clarity, we describe some examples of missingness. An Australian football data collector was taking lunch during the third quarter, so no data was recorded for this time period. Compare this to an AFL game being cancelled due to COVID19. There is no missing data because there is no data to be observed. If a study screens out individuals who do not meet inclusion criteria, this is not a missing value, because there is no intent for these values to be observed or measured.
